---
title: "R Notebook"
output:

---


## Introduction



Trying various interpolation methods IDW/kriging at one point in time

```{r}
rm(list = ls(all.names = TRUE))
```

```{r}
#base packages
library(tidyverse)
library(sf) 
library(sp) 
library(raster)

# geostatistics packages 
library(gstat) 
library(automap)

#plotting packages
library(patchwork)
library(viridis)

library(scales)
```

what i want is to pick a few hours on one day where the fires started to get bad
I want to create a variogram at each time point
To do this l'l need n dataframes, one for each hour

Each dataframe will have an x coord, y coord, PM10 value (maybe PM2.5 or another)


Import data
```{r}
load("C:/Users/Declan/Documents/STDS Assignment 3/AQI Data/working_databse/Combine_data_AQI_SWOQ_PM10_2019_2020_hourly.Rda")

load("C:/Users/Declan/Documents/STDS Assignment 3/AQI Data/working_databse/station_details_added_SWOQ.Rda")

load("C:/Users/Declan/Documents/STDS Assignment 3/AQI Data/working_databse/station_details_added_SWOQ.Rda")
```

Select Date 
```{r}
chosen_date<-subset(Combine_data_AQI_SWOQ_PM10_2019_2020_hourly, Date=="2019-12-25")
chosen_date<-subset(chosen_date, Hour==12)
```

Change date time
```{r}
chosen_date$date_time <- as.POSIXct(paste(chosen_date$Date, chosen_date$Hour),format = "%Y-%m-%d%H", tz='GMT')

chosen_date<- subset(chosen_date, select = -c(Date, Hour))
```

spread to do a count sum - don't select stations will more than 2 missing
```{r}
chosen_date_spread<- subset(chosen_date,select = -c(Longitude, Latitude,Region)) 
chosen_date_spread <- spread(chosen_date_spread,SiteName, Value)

chosen_date_spread<-chosen_date_spread[, colSums(is.na(chosen_date_spread)) <= 0]
```

impute missing values after rounding to 0
```{r}
chosen_date_spread[,-1][chosen_date_spread[,-1] <=0] <- 0    
#chosen_date_spread_imp <- na_kalman(chosen_date_spread)
#chosen_date_spread_imp<-chosen_date_spread
```

```{r}
station_value <- gather(chosen_date_spread, date_time) # ordered by hour
```

Change col names
```{r}
colnames(station_value) <- c("SiteName", "PM10")
```

```{r}
station_details<-subset(station_details, select = c(SiteName,Longitude,Latitude))
```

```{r}
hour_1_final <- left_join(station_value, station_details, by = "SiteName")
hour_1_final <- hour_1_final[,-1]
```


Following this tutorial

https://stackoverflow.com/questions/51501256/plotting-empirical-and-fitted-semivariogram-in-ggplot

Setting this up is the issue
Adding more here---
```{r}
hour_1_final_tibble<-as.tibble(hour_1_final)
```

```{r}
hour_1_final_sf <- st_as_sf(hour_1_final_tibble, coords = c("Longitude", "Latitude"), crs = "") %>% # testcrs = '+init=epsg:4326') %>% 
  cbind(st_coordinates(.))
```

```{r}
# We will discuss later, what Z~1 does actually mean in this context
my_v_emp_OK <- gstat::variogram(
  PM10~1,
  as(hour_1_final_sf, "Spatial") # switch from {sf} to {sp}
  )


plot(my_v_emp_OK)
```

```{r}
#vgm1.fit = fit.variogram(my_v_emp_OK, model = vgm(1, "Sph", 900, 1))
#vgm1.fit


my_v_emp_OK.fit = fit.variogram(my_v_emp_OK, model = vgm(1, "Sph", 0.15, 40))
my_v_emp_OK.fit

plot(my_v_emp_OK, my_v_emp_OK.fit)
```

```{r}
# automap's autofitVariogram actually produces more info than we need.
# I will only keep the var_model part.
my_v_mod_OK <- automap::autofitVariogram(PM10~1, as(hour_1_final_sf, "Spatial"))$var_model

# To inspect the automatic fit that was chosen for us we can use
# automap's excellent build in methods for base::plot
plot(automap::autofitVariogram(PM10~1, as(hour_1_final_sf, "Spatial")))
```

Can also find variogram now that cooridnates are sp

```{r}
hour_1_final_grd_100_sf <- hour_1_final_sf %>% 
  st_bbox() %>% 
  st_as_sfc() %>% 
  st_make_grid(
  cellsize = c(0.005, 0.005), # 100m pixel size
  what = "centers"
  ) %>%
  st_as_sf() %>%
  cbind(., st_coordinates(.))

hour_1_final_grd_100_sp <- as(hour_1_final_grd_100_sf, "Spatial") # converting to {sp} format
gridded(hour_1_final_grd_100_sp) <- TRUE             # informing the object that it is a grid
hour_1_final_grd_100_sp <- as(hour_1_final_grd_100_sp, "SpatialPixels") # specifying what kind of grid
```

```{r}
ggplot(data =hour_1_final_grd_100_sf, aes(x = X, y = Y))+
geom_point(size = 0.5)
```



```{r}
# Ordinary Kriging
my_OK_hour <- krige(
  PM10~1,                       # Z is our variable and "~1" means "depends on mean"
  as(hour_1_final_sf, "Spatial"), # input data in {sp} format
  hour_1_final_grd_100_sp,                # locations to interpolate at
  model = my_v_mod_OK           # the variogram model fitted above
  )
```

```{r}
my_OK_hour %>% as.data.frame %>%
  ggplot(aes(x=coords.x1, y=coords.x2)) + geom_tile(aes(fill=var1.pred)) + coord_equal() +
  scale_fill_gradient(low = "yellow", high="red") +
  scale_x_continuous(labels=comma) + scale_y_continuous(labels=comma) +
  theme_bw()
```

```{r}
x <- krige.cv(PM10~1, as(hour_1_final_sf, "Spatial"), my_v_mod_OK, nmax = 100, nfold=17)
bubble(x, "residual", main = "PM10: Leave on one residuals")

```
```{r}
kriging_result = autoKrige(PM10~1, as(hour_1_final_sf, "Spatial"), hour_1_final_grd_100_sp)
```
```{r}
plot(kriging_result)
```



```{r}
#Universal Kriging
my_UK_hour <- krige(
  PM10~coords.x1+coords.x2, # Think "Z~X+Y" but {sp} conversion alters variable naming
  as(hour_1_final_sf, "Spatial"), # input data in {sp} format (`X` --> `coords.x1`)
  hour_1_final_grd_100_sp,                # locations to interpolate at
  model = autofitVariogram(  # we need an appropriate variogram fit
    PM10~X+Y,                   # here we can keep "X+Y" - it's just how it is
    as(hour_1_final_sf, "Spatial")
    )$var_model
  )
```

```{r}
my_UK_hour %>% as.data.frame %>%
  ggplot(aes(x=coords.x1, y=coords.x2)) + geom_tile(aes(fill=var1.pred)) + coord_equal() +
  scale_fill_gradient(low = "yellow", high="red") +
  scale_x_continuous(labels=comma) + scale_y_continuous(labels=comma) +
  theme_bw()
```
```{r}
x <- krige.cv(PM10~1, as(hour_1_final_sf, "Spatial"), model=my_UK_hour,  nfold=17)
bubble(x, "residual", main = "PM10: Leave one out residuals")
```





```{r}
my_idwres_hour <- idw(
  PM10~1,                       # idw also depends on mean
  as(hour_1_final_sf, "Spatial"), # input data in {sp} format
  hour_1_final_grd_100_sp,                # locations to interpolate at
) 
```

```{r}
my_idwres_hour %>% as.data.frame %>%
  ggplot(aes(x=coords.x1, y=coords.x2)) + geom_tile(aes(fill=var1.pred)) + coord_equal() +
  scale_fill_gradient(low = "yellow", high="red") +
  scale_x_continuous(labels=comma) + scale_y_continuous(labels=comma) +
  theme_bw()
```

```{r}
df_hour_1_final <-as.data.frame(hour_1_final)
```


```{r}
p_hour_1_final_hour <- ggplot(
  data = df_hour_1_final,
  mapping = aes(x = Latitude, y = Longitude, color = PM10)
) +
  geom_point(size = 3) + 
  scale_color_viridis(option = "B",  limits = c(0, 20)) +
  ggtitle(label = "Observation PM10 across various stations") +
  theme_void() +
    theme(
      plot.title = element_text(hjust = 0.5)
    )

p_hour_1_final_hour
```

```{r}
library(ozmaps)

sydney_map <- ozmaps::abs_ced %>% filter(NAME %in% c(
  "Sydney", "Wentworth", "Warringah", "Kingsford Smith", "Grayndler", "Lowe", 
  "North Sydney", "Barton", "Bradfield", "Banks", "Blaxland", "Reid", 
  "Watson", "Fowler", "Werriwa", "Prospect", "Parramatta", "Bennelong", 
  "Mackellar", "Greenway", "Mitchell", "Chifley", "McMahon"
))


ggplot(sydney_map) + 
  geom_sf(aes(fill = NAME), show.legend = FALSE) + 
  coord_sf(xlim = c(150.97, 151.3), ylim = c(-33.98, -33.79)) 
```

```{r}
Sydney_map <- ozmaps::abs_ced 

#xmin: ymin:  xmax:  ymax: 

p <- ggplot(Sydney_map) + geom_sf()
p + coord_sf(expand = FALSE, xlim = c(150.5007 , 151.301), ylim = c(-34.30371, -33.61871))111
```
```{r}
library(leaflet)

#aqi_bounds_lat1 <-  floor(aqi_lat_range[1])
#aqi_bounds_long1 <- floor(aqi_long_range[1])
#aqi_bounds_lat2 <- ceiling(aqi_lat_range[2])
#aqi_bounds_long2 <- ceiling(aqi_long_range[2])
# adding markers`S

aqi_map <- leaflet() %>%
  
  fitBounds(150.5007,-34.30371,151.2407,-33.61871 )%>%

  addTiles() %>%
  addMarkers(station_details$Longitude, 
             station_details$Latitude, 
             popup = station_details$SiteName) 


aqi_map
```



