---
title: "R Notebook"
output: html_notebook
---
# Trying to get spacetime to work again

```{r}
rm(list = ls(all.names = TRUE))
```

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(lubridate)
library(leaflet)
library(gdata)
library(imputeTS)
library(magrittr)
library(stringr)
library(thief)
library(forecast)
library(rgdal)
library(zoo)
library(reshape2)
```
Import data
```{r}
load("C:/Users/Declan/Documents/STDS Assignment 3/AQI Data/working_databse/Air Quality Site Details.Rda")

load("C:/Users/Declan/Documents/DataScience/Air pollution copy before redownloading/AQI_2016_2020_condensed_v2.Rda") 
```

```{r}
working_aqi_dataset = merge(AQI_2016_2020_work,df_site_details, by='Site_Id')

rm("AQI_2016_2020_work")
```

Removing columns that aren't interesting
```{r}
working_aqi_dataset <- subset(working_aqi_dataset,                               select=-c(Site_Id,AirQualityCategory, Parameter.ParameterDescription, Parameter.UnitsDescription))
```

```{r}
#unique(working_aqi_dataset$Region)
```

```{r}
#unique(working_aqi_dataset$SiteName)
```

Select only PM 10
using GMT as timezone
```{r}
PM10<-subset(working_aqi_dataset, Parameter.ParameterCode=="PM10")

PM10 <- PM10[grep("Sydney East|Sydney South-west|Central Coast|Lower Hunter|Upper Hunter|Roadside Monitoring|Sydney North-west|Newcastle Local|Illawarra|Research Monitoring|Central Tablelands",PM10$Region), ]
           

PM10$date_time <- as.POSIXct(paste(PM10$Date, PM10$Hour),format = "%Y-%m-%d%H", tz='GMT')

PM10<-subset(PM10, select = -c(Longitude, Latitude,  Parameter.ParameterCode))
```

```{r}
rm(working_aqi_dataset)
```


spreading the values
```{r}
PM10_spread <- subset(PM10, select=-c(Region))
PM10_spread <- spread(PM10_spread,SiteName, Value)
```

```{r}
rm(PM10)
```

Counting column values
Total of 24*365.25*5 = 43830 values 



```{r}
PM10_spread<-PM10_spread[, colSums(is.na(PM10_spread)) <= 1000]
```



#drop first 2 colums date and hour
```{r}
PM10_spread <-PM10_spread[,-1:-2] # hour date
```

Fixing values below 0 and rounding 
```{r}
#PM10_spread_dt<-PM10_spread[,1]
PM10_spread[,-1][PM10_spread[,-1] <=1] <- 1 # returns  0 after taking the log   
#PM10_spread[,-1]<-log(PM10_spread[,-1]) ->>>>>>>
PM10_spread[,-1] <-round(PM10_spread[,-1],4)
```

imputing missing values
```{r}
PM10_imp <- na_kalman(PM10_spread)
```

```{r}
names(PM10_imp) <- gsub(" ", "_", names(PM10_imp))
```

```{r}
rm(PM10_spread)
```


```{r}
#colnames(PM10_imp)
```

```{r}
#qplot(x=final_time, y=PROSPECT, data = PM10_imp)
```
As there are numerous missing values in PM2.5, it is going to be dropped. PM10 will include all PM2.5

```{r}

#unzip("C:/Users/Declan/Documents/STDS Assignment 3/nsw_lga_polygon_shp.zip")
```

```{r}
#NSWLGA <- readOGR(dsn = "NSW_LGA_POLYGON_shp", layer = "NSW_LGA_POLYGON_shp")
```

Spatial Temporal stuff
```{r}
library(spacetime)
library(gstat)
library(automap)
library(rgdal)
library(rgeos)
library(sp)
#library(raster)
```

Select only 2019 dates and reindex
```{r}
PM10_subset_2019 <- tail(PM10_imp,n = 43848-26304)
rownames(PM10_subset_2019) <- 1:nrow(PM10_subset_2019)
PM10_subset_2019<- head(PM10_subset_2019,n = 8760)
```

Select last year - 2020 and reindex
```{r}
PM10_subset_2020 <- tail(PM10_imp,n = 43848-35064)
rownames(PM10_subset_2020) <- 1:nrow(PM10_subset_2020)
```

```{r}
#sapply(PM10_subset_2019, max)
```


```{r}
subset_df_sites <- df_site_details

subset_df_sites <- subset_df_sites[grep("ABERDEEN|ALBION PARK STH|BARGO|BERESFIELD|BULGA|CAMBERWELL|CAMPBELLTOWN WEST|CARRINGTON|CHULLORA|JERRYS PLAINS|LIVERPOOL|MAISON DIEU|MAYFIELD|MERRIWA|MUSWELLBROOK|MUSWELLBROOK NW|NEWCASTLE|OAKDALE|PROSPECT|RANDWICK|SINGLETON|SINGLETON NW|SINGLETON SOUTH|STOCKTON|WARKWORTH|WYBONG|WYONG",subset_df_sites$SiteName), ]
```

```{r}
coordinates(subset_df_sites)<- ~ Longitude + Latitude
```

```{r}
proj4string(subset_df_sites) <- CRS("+proj=longlat +ellps=WGS84")
```
No missing values
```{r}
#colSums(is.na(PM10_subset_2020)) 
```


```{r}
PM10_2019_zoo <-zoo(PM10_subset_2019[,-1], PM10_subset_2019$date_time)
PM10_2020_zoo <-zoo(PM10_subset_2020[,-1], PM10_subset_2020$date_time) 
```

```{r}
PM10_2019 <- data.frame(vals = as.vector(t(PM10_2019_zoo)))  
PM10_2020 <- data.frame(vals = as.vector(t(PM10_2020_zoo))) 
```

```{r}
  PM10_2019_st <- STFDF(sp = subset_df_sites,                   
                time = index(PM10_2019_zoo),
                data = PM10_2019)

PM10_2020_st <- STFDF(sp = subset_df_sites,                   
                time = index(PM10_2020_zoo),
                data = PM10_2020)
```


Spacetime graphics
```{r}
airPal <- colorRampPalette(c('springgreen1', 'sienna3', 'gray5'))(5)
  
stplot(PM10_2019_st[, 1:2],
       cuts = 5,
       col.regions = airPal,
       main = '',
       edge.col = 'black')

stplot(PM10_2019_st, mode = 'xt',
       col.regions = colorRampPalette(airPal)(15),
       scales = list(x = list(rot = 45)),
       ylab = '', xlab = '', main = '')

stplot(PM10_2019_st, mode = 'ts',
       xlab = '',
       lwd = 0.1, col = 'black', alpha = 0.6,
       auto.key = FALSE)
```

variogram 
```{r}
vario_2019 <- variogramST(vals~1, 
data=PM10_2019_st, assumeRegular=T)

vario_2020 <- variogramST(vals~1, 
data=PM10_2020_st)


#tunit = "hours",
#assumeRegular=F #PM10_2019
```


page 50 
https://reseau-resste.mathnum.inrae.fr//sites/default/files/2020-09/analyzing-spatio-temporal-data.pdf
```{r}
variot <- function(model,vario=vario_2019){
tseq <- seq(0,400,20)
grid <- expand.grid(seq(0,50000,10),tseq)
dg <- data.frame(spacelag=grid[,1],timelag=grid[,2])
table<- variogramSurface(model,dg)
nx <- length(seq(0,50000,10))
ny <- length(seq(0,15,1))
plot(vario[vario$dist==0,5],vario[vario$dist==0,3],ylim=range(c(0,vario[vario$dist==0,3],
table[table[,1]==0,3]),na.rm=T))
lines(tseq,table[table[,1]==0,3])
}
```

```{r}
separableModel <- vgmST("separable",space=vgm(0.9,"Exp",100,0.1),
time=vgm(0.9,"Exp",3.5,0.1),sill=40)

separable_fit <- fit.StVariogram(model=separableModel,object=vario_2019)

plot(vario,separable_fit,all=T,map=F)

```

```{r}
sill <- 800
separable_man <- vgmST("separable",
                       space=vgm(3,"Exp",100,0.01),
                       time=vgm(1/sill,"Exp",100,19/sill),
                       sill=sill)

plot(vario,separable_man,all=F,map=F)

variot(separableModel,vario_2019)
```

```{r}

```
```{r}
ProductSum <- vgmST("productSum",space =vgm(9,"Exp",8e3,1),time=vgm(8,"Exp",106,2),k=2)
ProductSum_fit <- fit.StVariogram(model=ProductSum,object=vario)
plot(vario,ProductSum_fit,all=T,map=F)

```
```{r}
variot(ProductSum,vario)

```

